---
title: "Supplementary Material 2 - Model Selection Procedure for Correlates of Burrow Microclimate and Buffering Capacity"
subtitle: "Burrows buffer nest temperatures and offer a stable thermal microclimate for threatened seabird chicks during extreme events"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}

lssp <- read.csv("lssp.csv")

lssp$Burrow <- factor(lssp$Burrow)

library(DHARMa); library(lme4); library(mgcViz); library(mgcv);library(dsm);library(dplyr)

```


# PETREL MODEL - correlates of burrow microclimate

Here we show the steps followed to test for correlation between external environmental conditions and petrel burrow thermal environment.

We tested a variety of models including GAM and GAMM. Additionally we tried different methods of accounting for autocorrelation. 


## Model S2.1, S2.2 and S2.3 - Initial Models

Run three initial models to identify whether the weather variables should be fitted as spline terms, a spline interaction term or linear interaction term. 

```{r, warning = FALSE, message=FALSE}

# MODEL S2.1
# Model with spline fits
m1_petrel <- gam(data = lssp,
           Inside~s(scale(Temperature), bs="ts") + # Weather 
                  s(scale(Wind.Speed),bs="ts") +  # Weather
                  s(Time, bs="cc") + # Time
                  s(julian, bs="ts") + # Date
                  s(Burrow, bs="re"), # random effect
                  method = "REML") 

# MODEL S2.2
# Explore temp-wind interaction term
m2_petrel <- gam(data = lssp,
           Inside~s(scale(Temperature), scale(Wind.Speed), bs="ts") +
                  s(Time, bs="cc") + # Time
                  s(julian, bs="ts") + # Date
                  s(Burrow, bs="re"), # random effect
                  method = "REML") 

# MODEL S2.3
# Explore Linear interaction 
m3_petrel <- gam(data = lssp,
           Inside~scale(Temperature) * scale(Wind.Speed) + 
                  s(Time, bs="cc") + # Time
                  s(julian, bs="ts") + # Date
                  s(Burrow, bs="re"), # random effect
                  method = "REML") 

```

**Compare AIC**

Spline interaction (`m2_petrel`) has the lowest AIC score.

```{r}
AIC(m1_petrel, m2_petrel, m3_petrel)
```



**Model Summary**
```{r, echo=FALSE}

summary(m2_petrel)

```

## Model S2.4 - Increase the knots for Julian date

From the summary table of `m2_petrel`, the edf is close to the ref.df, indicating that the knots may need to be increased. Here we double the knots to 20.

```{r, warning = FALSE, message=FALSE}

# MODEL S2.4
m4_petrel <- gam(data = lssp,
           Inside~s(scale(Temperature), scale(Wind.Speed), bs="ts") +
                  s(Time, bs="cc") + # Time
                  s(julian, bs="ts", k = 20) + # Date
                  s(Burrow, bs="re"), # random effect
                  method = "REML") 

```


**Model Summary**
```{r, echo=FALSE}

summary(m4_petrel)

```

## Model S2.5 - Increase the knots again for Julian date
```{r, warning = FALSE, message=FALSE}

# MODEL S2.5
m5_petrel <- gam(data = lssp,
           Inside~s(scale(Temperature), scale(Wind.Speed)) +
                  s(Time, bs="cc") + # Time
                  s(julian, bs="ts", k = 40) + # Date
                  s(Burrow, bs="re"), # random effect
                  method = "REML") 

```

**Model Summary**

From the summary of `m5_petrel`, the edf continues to be close to the ref.df value, but we are unable to increase the knots further.

```{r, echo=FALSE}

summary(m5_petrel)

```

**Compare AIC**

Model `m5_petrel` with k = 40 has the lowest AIC score. 

```{r}
AIC(m2_petrel, m4_petrel, m5_petrel)
```


## Explore model fit

**Test for temporal autocorrelation**

The model `m5_petrel` is highly autocorrelated, therefore, we reject this model.
```{r, out.width = '60%', fig.align = "center"}
acf(residuals(m5_petrel))
```


## Model S2.6 - `GAMM` with corAR1

We will also explore the GAMM option with burrow as a random effect and include a corAR1 term to account for autocorrelation. We removed the date and time terms because the corAR1 term is accounting for the autocorrelation. 

```{r}

# MODEL S2.6
gamm1_petrel <- gamm(data = lssp,
                   Inside~s(scale(Temperature), scale(Wind.Speed)),
                   random = list(Burrow = ~1),# Burrow features
                   corr = corAR1(form = ~ 1|Burrow),
                   method = "REML")

```

**Model Summary - gam**
```{r, echo=FALSE}

summary(gamm1_petrel$gam)

```

**Model Summary - lme**
```{r, echo=FALSE}

summary(gamm1_petrel$lme)

```

## Explore model fit

**Test for temporal autocorrelation**

The residual function for gam in R returns the residuals without adjustment for the correlation structure. So must specify type = "normalized"
```{r, out.width = '60%', fig.align = "center"}
acf(residuals(gamm1_petrel$lme, type = "normalized"))
```


**Gam.check**

```{r, fig.height = 15, fig.width = 15, fig.align = "center"}

gam.check(gamm1_petrel$gam)

```



**lme plot**

``````{r, out.width = '60%', fig.align = "center"}

plot(gamm1_petrel$lme)

```



**Visualise the temp-wind interaction in 2D**

``````{r, out.width = '60%', fig.align = "center"}

plot(gamm1_petrel$gam, scheme = 2, pages = 1)

```



# PUFFIN MODEL - correlates of burrow microclimate

Here we show the modelling steps followed to arrive at the final selected model used to test correlates of Atlantic puffin burrow microclimates.  


```{r, include=FALSE}
atpu <- read.csv("atpu.csv")

atpu$Burrow <- factor(atpu$Burrow)

```


## Exploring wind speed

In the plot below, the data for wind speed exceed the speeds recorded during the hurricane, suggesting that there may be issues with data at these three points. They occurred all on the same day over a 1.5 hour period. 

```{r, out.width = '60%', fig.align = "center"}
plot(atpu$Wind.Speed, atpu$Inside)
```

**Removing potentially erroneous data**

We will remove the high wind speed records for the modelling.

```{r, out.width = '60%', fig.align = "center"}
# Remove the higher wind speeds
atpu.wind <- atpu %>% filter(Wind.Speed < 30)

# Visualise the data
plot(atpu.wind$Wind.Speed, atpu.wind$Inside)
```


## Model S2.7, S2.8, and S2.9 - Initial models

Run three initial models to identify whether the weather variables should be fitted as spline terms, a spline interaction term or linear interaction term. 

```{r, warning = FALSE, message=FALSE}

# MODEL S2.7
# Model with spline fits
m1_puffin <- gam(data = atpu.wind,
           Inside~s(scale(Temperature), bs="ts") + # Weather 
                  s(scale(Wind.Speed),bs="ts") +  # Weather
                  s(Time, bs="cc") + # Time
                  s(julian, bs="ts") + # Date
                  s(Burrow, bs="re"), # random effect
                  method = "REML") 

# MODEL S2.8
# Explore temp-wind interaction term
m2_puffin <- gam(data = atpu.wind,
           Inside~s(scale(Temperature), scale(Wind.Speed)) +
                  s(Time, bs="cc") + # Time
                  s(julian, bs="ts") + # Date
                  s(Burrow, bs="re"), # random effect
                  method = "REML") 

# MODEL S2.9
# Explore Linear interaction 
m3_puffin <- gam(data = atpu.wind,
           Inside~scale(Temperature) * scale(Wind.Speed) + 
                  s(Time, bs="cc") + # Time
                  s(julian, bs="ts") + # Date
                  s(Burrow, bs="re"), # random effect
                  method = "REML") 


```

**Compare AIC**

Spline interaction (`m2_puffin`) has the lowest AIC score

```{r}
AIC(m1_puffin, m2_puffin, m3_puffin)
```



**Model Summary**
```{r, echo=FALSE}

summary(m2_puffin)

```


## Model S2.10 - Increase the knots for Julian date

From the summary table of `m2_puffin`, the edf is close to the ref.df, indicating that the knots may need to be increased. Here we double the knots to 20. 

```{r}
# MODEL S2.10
m4_puffin <- gam(data = atpu.wind,
             Inside~s(scale(Temperature), scale(Wind.Speed)) +
                  s(Time, bs="cc") + 
                  s(julian, bs="ts", k = 20)+ # Time and date
                  s(Burrow, bs="re"), # random effect
                  method = "REML") 
```


**Model Summary**

The edf continues to be close to the ref.df value, but we are unable to increase the knots further. 

```{r, echo=FALSE}

summary(m4_puffin)

```

**Compare AIC**

Increasing the knots reduces the AIC score. 

```{r}
AIC(m2_puffin, m4_puffin)
```


## Model S2.11 - Increase the knots for the interaction term

From the summary table of `m4_puffin`, the edf is close to the ref.df, indicating that the knots may need to be increased. Here we double to knots to 60. 

```{r}
# MODEL S2.11
m5_puffin <- gam(data = atpu.wind,
             Inside~s(scale(Temperature), scale(Wind.Speed), k = 60) +
                  s(Time, bs="cc") + 
                  s(julian, bs="ts", k = 20)+ # Time and date
                  s(Burrow, bs="re"), # random effect
                  method = "REML") 
```


**Model Summary**
```{r, echo=FALSE}

summary(m5_puffin)

```

**Compare AIC**

Increasing the knots reduces the AIC score. We will now explore the model fit.

```{r}
AIC(m4_puffin, m5_puffin)
```


## Explore model fit

**Test for temporal autocorrelation**

The model `m5_puffin` is highly autocorrelated, therefore, we reject this model. 

```{r, out.width = '60%', fig.align = "center"}
acf(residuals(m5_puffin))
```


## Model S2.12 - `GAMM` with corAR1

We will also explore the GAMM option with burrow as a random effect and include a corAR1 term to account for autocorrelation. We removed the date and time terms because the corAR1 term is accounting for the autocorrelation. 

```{r}
## Model S2.12
gamm1_puffin <- gamm(data = atpu.wind,
                   Inside~s(scale(Temperature), scale(Wind.Speed)),
                   random = list(Burrow = ~1),# Burrow features
                   corr = corAR1(form = ~ 1|Burrow),
                   method = "REML")

```

**Model Summary - gam**
```{r, echo=FALSE}

summary(gamm1_puffin$gam)

```

**Model Summary - lme**
```{r, echo=FALSE}

summary(gamm1_puffin$lme)

```

## Explore model fit

**Test for temporal autocorrelation**

The residual function for gam in R returns the residuals without adjustment for the correlation structure. So must specify type = "normalized"
```{r, out.width = '60%', fig.align = "center"}
acf(residuals(gamm1_puffin$lme, type = "normalized"))
```


**Gam.check**

```{r, fig.height = 15, fig.width = 15, fig.align = "center"}

gam.check(gamm1_puffin$gam)

```



**lme plot**

``````{r, out.width = '60%', fig.align = "center"}

plot(gamm1_puffin$lme)

```



**Visualise the temp-wind interaction in 2D**

``````{r, out.width = '60%', fig.align = "center"}

plot(gamm1_puffin$gam, scheme = 2, pages = 1)

```




# PETREL MODELS - Favourable burrow features during extreme temperatures 

Filter the data in the 1st and 99th percentiles of external air temperatures recorded at the weather stations. These temperatures were identified as 25.7째C and 6.5째C.

```{r}

## HOT
extreme_heat_LSSP <- filter(lssp, Temperature >= 25.7 & Species == "Leach's storm-petrel")
extreme_heat_LSSP$Extreme <- "Heat Extreme"


## COLD
extreme_cold_LSSP <- filter(lssp, Temperature <= 6.5 & Species == "Leach's storm-petrel")
extreme_cold_LSSP$Extreme <- "Cold Extreme"

```


## Model S2.13 - `GAM` model - Heat extreme

```{r}
# MODEL S2.13
gam_heat_lssp<- gam(data = extreme_heat_LSSP,
           buffer_kestrel~s(scale(Canopy), bs="ts") + # Weather 
                  s(scale(Enterance_Area),bs="ts") +  # Weather
                  s(scale(volume),bs="ts") +  # Weather
                  s(Burrow, bs="re"), # random effect
                  method = "REML") 

```


**Model Summary**
```{r, echo=FALSE}

summary(gam_heat_lssp)

```


## Explore model fit

**Test for temporal autocorrelation**

```{r, out.width = '60%', fig.align = "center"}
acf(residuals(gam_heat_lssp))
```


**Gam.check**

```{r, fig.height = 15, fig.width = 15, fig.align = "center"}
gam.check(gam_heat_lssp)
```


**Plotting**

```{r, fig.height = 15, fig.width = 15, fig.align = "center"}
plot(gam_heat_lssp, pages = 1)
```


## Model S2.14 - `GAMM` with corAR1 - Heat extreme

```{r}

# MODEL S2.14
gamm_heat_lssp <- gamm(data = extreme_heat_LSSP,
                   buffer_kestrel~ s(scale(Canopy), bs="ts") +
                   s(scale(Enterance_Area),bs="ts") +  # 
                   s(scale(volume),bs="ts"),  # Weather
                   random = list(Burrow = ~1),# Burrow features
                   corr = corAR1(form = ~ 1|Burrow),
                   method = "REML")

```


**Model Summary - gam**
```{r, echo=FALSE}

summary(gamm_heat_lssp$gam)

```

**Model Summary - lme**
```{r, echo=FALSE}

summary(gamm_heat_lssp$lme)

```

## Explore model fit

**Test for temporal autocorrelation**

The residual function for gam in R returns the residuals without adjustment for the correlation structure. So must specify type = "normalized"

```{r, out.width = '60%', fig.align = "center"}
acf(residuals(gamm_heat_lssp$lme, type = "normalized"))
```


**Gam.check**

```{r, fig.height = 15, fig.width = 15, fig.align = "center"}

gam.check(gamm_heat_lssp$gam)

```


**lme plot**

``````{r, out.width = '60%', fig.align = "center"}

plot(gamm_heat_lssp$lme)

```


**Plot**

``````{r, fig.height = 15, fig.width = 15, fig.align = "center"}

plot(gamm_heat_lssp$gam, pages = 1)

```



## Model S2.15 - `GAM` model - Cold extreme

```{r}
# MODEL S2.15
gam_cold_lssp<- gam(data = extreme_cold_LSSP,
           buffer_kestrel~s(scale(Canopy), bs="ts") + # Weather 
                  s(scale(Enterance_Area),bs="ts") +  # Weather
                  s(scale(volume),bs="ts") +  # Weather
                  s(Burrow, bs="re"), # random effect
                  method = "REML") 

```


**Model Summary**
```{r, echo=FALSE}

summary(gam_cold_lssp)

```


## Explore model fit

**Test for temporal autocorrelation**

```{r, out.width = '60%', fig.align = "center"}
acf(residuals(gam_cold_lssp))
```


**Gam.check**

```{r, fig.height = 15, fig.width = 15, fig.align = "center"}
gam.check(gam_cold_lssp)
```


**Plotting**

```{r, fig.height = 15, fig.width = 15, fig.align = "center"}
plot(gam_cold_lssp, pages = 1)
```


## Model S2.16 - `GAMM` with corAR1 - cold extreme

```{r}
# MODEL S2.16
gamm_cold_lssp <- gamm(data = extreme_cold_LSSP,
                   buffer_kestrel~ s(scale(Canopy), bs="ts") +
                   s(scale(Enterance_Area),bs="ts") +  # 
                   s(scale(volume),bs="ts"),  # Weather
                   random = list(Burrow = ~1),# Burrow features
                   corr = corAR1(form = ~ 1|Burrow),
                   method = "REML")

```


**Model Summary - gam**
```{r, echo=FALSE}

summary(gamm_cold_lssp$gam)

```

**Model Summary - lme**
```{r, echo=FALSE}

summary(gamm_cold_lssp$lme)

```

## Explore model fit

**Test for temporal autocorrelation**

The residual function for gam in R returns the residuals without adjustment for the correlation structure. So must specify type = "normalized"

```{r, out.width = '60%', fig.align = "center"}
acf(residuals(gamm_cold_lssp$lme, type = "normalized"))
```


**Gam.check**

```{r, fig.height = 15, fig.width = 15, fig.align = "center"}

gam.check(gamm_cold_lssp$gam)

```



**lme plot**

``````{r, out.width = '60%', fig.align = "center"}

plot(gamm_cold_lssp$lme)

```


**Plot**

``````{r, fig.height = 15, fig.width = 15, fig.align = "center"}

plot(gamm_cold_lssp$gam, pages = 1)

```





# PUFFIN MODELS - Favourable burrow features during extreme temperatures 

Filter the data in the 1st and 99th percentiles of external air temperatures recorded at the weather stations. These temperatures were identified as 25.6째C and 10째C.

```{r}

## COLD
extreme_cold_ATPU <- filter(atpu.wind, Temperature <= 10 & Species == "Atlantic puffin")
extreme_cold_ATPU$Extreme <- "Cold Extreme"

## HOT
extreme_heat_ATPU <- filter(atpu.wind, Temperature >= 25.6 & Species == "Atlantic puffin")
extreme_heat_ATPU$Extreme <- "Heat Extreme"

```


## Model S2.17 - `GAM` model - Heat extreme

```{r}
# MODEL S2.17
gam_heat_ATPU<- gam(data = extreme_heat_ATPU,
           buffer_kestrel~s(scale(Enterance_Area),bs="ts") +  # Weather
                  s(scale(volume),bs="ts") +  # Weather
                  s(Burrow, bs="re"), # random effect
                  method = "REML") 

```


**Model Summary**
```{r, echo=FALSE}

summary(gam_heat_ATPU)

```


## Explore model fit

**Test for temporal autocorrelation**

```{r, out.width = '60%', fig.align = "center"}
acf(residuals(gam_heat_ATPU))
```


**Gam.check**

```{r, fig.height = 15, fig.width = 15, fig.align = "center"}
gam.check(gam_heat_ATPU)
```


**Plotting**

```{r, fig.height = 15, fig.width = 15, fig.align = "center"}
plot(gam_heat_ATPU, pages = 1)
```


## Model S2.18 - `GAMM` with corAR1 - Heat extreme

```{r}
# MODEL S2.18
gamm_heat_ATPU <- gamm(data = extreme_heat_ATPU,
                   buffer_kestrel~ s(scale(Enterance_Area),bs="ts") +  # 
                   s(scale(volume),bs="ts"),  # Weather
                   random = list(Burrow = ~1),# Burrow features
                   corr = corAR1(form = ~ 1|Burrow),
                   method = "REML")

```


**Model Summary - gam**
```{r, echo=FALSE}

summary(gamm_heat_ATPU$gam)

```

**Model Summary - lme**
```{r, echo=FALSE}

summary(gamm_heat_ATPU$lme)

```

## Explore model fit

**Test for temporal autocorrelation**

The residual function for gam in R returns the residuals without adjustment for the correlation structure. So must specify type = "normalized"

```{r, out.width = '60%', fig.align = "center"}
acf(residuals(gamm_heat_ATPU$lme, type = "normalized"))
```


**Gam.check**

```{r, fig.height = 15, fig.width = 15, fig.align = "center"}

gam.check(gamm_heat_ATPU$gam)

```


**lme plot**

``````{r, out.width = '60%', fig.align = "center"}

plot(gamm_heat_ATPU$lme)

```


**Plot**

``````{r, fig.height = 15, fig.width = 15, fig.align = "center"}

plot(gamm_heat_ATPU$gam, pages = 1)

```



## Model S2.19 - `GAM` model - Cold extreme

```{r}
## MODEL S2.19
gam_cold_ATPU<- gam(data = extreme_cold_ATPU,
           buffer_kestrel~s(scale(Enterance_Area),bs="ts") +  # Weather
                  s(scale(volume),bs="ts") +  # Weather
                  s(Burrow, bs="re"), # random effect
                  method = "REML") 

```


**Model Summary**
```{r, echo=FALSE}

summary(gam_cold_ATPU)

```


## Explore model fit

**Test for temporal autocorrelation**

```{r, out.width = '60%', fig.align = "center"}
acf(residuals(gam_cold_ATPU))
```


**Gam.check**

```{r, fig.height = 15, fig.width = 15, fig.align = "center"}
gam.check(gam_cold_ATPU)
```


**Plotting**

```{r, fig.height = 15, fig.width = 15, fig.align = "center"}
plot(gam_cold_ATPU, pages = 1)
```


## Model S2.20 - `GAMM` with corAR1 - cold extreme

```{r}
# Model S2.20
gamm_cold_ATPU <- gamm(data = extreme_cold_ATPU,
                   buffer_kestrel~ s(scale(Enterance_Area),bs="ts") +  # 
                   s(scale(volume),bs="ts"),  # Weather
                   random = list(Burrow = ~1),# Burrow features # random effect
                   corr = corAR1(form = ~ 1|Burrow),
                   method = "REML")

```


**Model Summary - gam**
```{r, echo=FALSE}

summary(gamm_cold_ATPU$gam)

```

**Model Summary - lme**
```{r, echo=FALSE}

summary(gamm_cold_ATPU$lme)

```

## Explore model fit

**Test for temporal autocorrelation**

The residual function for gam in R returns the residuals without adjustment for the correlation structure. So must specify type = "normalized"

```{r, out.width = '60%', fig.align = "center"}
acf(residuals(gamm_cold_ATPU$lme, type = "normalized"))
```


**Gam.check**

```{r, fig.height = 15, fig.width = 15, fig.align = "center"}

gam.check(gamm_cold_ATPU$gam)

```



**lme plot**

``````{r, out.width = '60%', fig.align = "center"}

plot(gamm_cold_ATPU$lme)

```


**Plot**

``````{r, fig.height = 15, fig.width = 15, fig.align = "center"}

plot(gamm_cold_ATPU$gam, pages = 1)

```
